|

### 2022-12-29 paper.tex
Tue Dec 27 03:05:17 CET 2022: update-remote.sh
autogit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

@@

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2022-12-23}

@@

DAE-GP builds upon the concept of using DAEs in EDAs described by
{[}11{]} and transfers the concept to the domain of GP. The mutation and
crossover operators of standard GP are replaced by sampling new
solutions from a probabilistic model that is build by training a DAEs
long short-term memory (LSTM) network on a subset of high fitness
solutions from the current population {[}18{]}. LSTMs are a variant of
recurrent neural networks first introduced by {[}7{]} that can store
learned information over en extended time period while avoiding the
problem of vanishing and/or exploding gradients. Since DAE-GP encodes
candidate solutions as linear strings in prefix notation, the DAE in
DAE-GP uses LSTMs for both encoding and decoding where the total amount
of time steps \(T\) is equal to sum of the length of the input solution
and the output solution {[}18{]}.

@@

The strategy of pre-training artificial neural networks is based on the
idea of transfer learning. Transfer learning in the field of machine
learning aims at retaining previously learned knowledge from one task to
re-use it for another task {[}5{]} {[}9{]}. Transfer learning
traditionally follows a two phase approach:

@@

where source task and target task are usually similar but may differ in
their feature space and the distribution of training data {[}9{]}. One
of the main motivations for using transfer learning in real world
machine learning tasks is the ability to reduce the need for large
amounts of training data which can often be unavailable or too expensive
to collect.

Pre-Training is a commonly used strategy in deep architectures that has
been shown to improve both the optimization process itself as well as
the generalization behavior if compared to randomly initialized
parameters {[}3{]}.

@@

The DAE-GP algorithm, first described by {[}18{]}, is described by
algorithm 1

\begin{Shaded}
\begin{Highlighting}[numbers=left,,]
\NormalTok{WHILE NOT termination\_criterion}
\NormalTok{  DO }
\NormalTok{END WHILE}
\NormalTok{RETURN}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
\caption{Airfoil - First Generation Median Training Error for variable
number of hidden Neurons}
\end{figure}

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
\caption{Airfoil - First Generation final Training Error for variable
number of hidden Neurons}
\end{figure}

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png}
\caption{Airfoil - First Generation Median Training Error for variable
number of hidden Layers}
\end{figure}

\begin{figure}
\centering
\includegraphics{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png}
\caption{Airfoil - First Generation final Training Error for variable
number of hidden Layers}
\end{figure}

@@

Regarding the hidden neurons, the dimension of each hidden layer, I used
two different strategies: Regular DAE-GP uses the same strategy as
described by earlier work (see {[}17{]} or {[}18{]}), the number of
hidden neurons is dynamically set at each generation to the maximum
length of all solutions inside the current training population. For
pre-trained DAE-GP another strategy had to be used, we use a fixed
number of hidden neurons for each run that is set equal to the maximum
length of all individuals that are present in the initial pre-training
population. I decided to take this aproach instead of using a fixed
number of hidden neurons for the fact that it simplifies the adjustment
of the parameter for later experiments on different problems. It should
nonetheless be mentioned that this strategy does give pre-training on
average more hidden neurons per layer resulting in a more complex
network (since the pre-training population is much larger than the
regular population, the chances for larger individuals inside it are
also higher).

\begin{figure}
\centering
\includegraphics{./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png}
\caption{Airfoil - Best Fitness over 30 Generations}
\end{figure}

\begin{figure}
\centering
\includegraphics{./img/airfoil_FullRun_30gens/final_fit_boxplot.png}
\caption{Airfoil - Fitness Distribution Boxplot}
\end{figure}

\includegraphics{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
Since {[}17{]} shows, that DAE-GP finds higher quality that are also
smaller in size than those found by average GP for a given number of
10.000 fitness evaluations, I also studied the differences in average
solution size between pre-trained and regular DAE-GP.

\begin{figure}
\centering
\includegraphics{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
\caption{Airfoil - Average Solution Size}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage[numbers]{natbib}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

Master Thesis - M.Sc. Business Education

@@

Johannes Gutenberg University Mainz}
\author{}
\date{\vspace{-2.5em}Date of Submission: 2022-12-29}

@@

DAE-GP builds upon the concept of using DAEs in EDAs described by
{[}11{]} and transfers the concept to the domain of GP. The mutation and
crossover operators of standard GP are replaced by sampling new
solutions from a probabilistic model that is build by training a DAEs
long short-term memory (LSTM) network on a subset of high fitness
solutions from the current population {[}18{]}.

LSTMs are a variant of recurrent neural networks first introduced by
{[}7{]} that can store learned information over en extended time period
while avoiding the problem of vanishing and/or exploding gradients.
Since DAE-GP encodes candidate solutions as linear strings in prefix
notation, the DAE in DAE-GP uses LSTMs for both encoding and decoding
where the total amount of time steps \(T\) is equal to sum of the length
of the input solution and the output solution {[}18{]}.

@@

Pre-Training describes the concept of initially training a neural
network on a large dataset before using it to solve a more specific
task. It is a commonly used strategy in deep architectures that has been
shown to improve both the optimization process itself as well as the
generalization behavior if compared to the standard approach of using
randomly initialized parameters {[}3{]}.

The strategy of pre-training artificial neural networks is based on the
idea of transfer learning that aims at retaining previously learned
knowledge from one task to re-use it for another task {[}5{]} {[}9{]}.
Transfer learning traditionally follows a two phase approach:

@@

where source task and target task are usually similar but may differ in
their feature space and the distribution of training data {[}9{]}.

Some of the main motivations for using pre-training is the ability to
reduce the need for large amounts of training data which can often be
unavailable or too expensive to collect as well as an improved
performance by either reducing the computational effort that is
necessary to train a model to solve a specific task or by improving the
models generalisation ability.

@@

The DAE-GP algorithm, first described by {[}18{]}, is summarized by
figure 1.

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/flowcharts/dae-gp.png}
\caption{Flowchart - Regular DAE-GP}
\end{figure}

The pre-training strategy implemented is visualized by figure 2.

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/flowcharts/pt-dae-gp.png}
\caption{Flowchart - Pre-Trained DAE-GP}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png}
\caption{First Generation Median Training Error for variable number of
hidden Neurons - Airfoil}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png}
\caption{First Generation final Training Error for variable number of
hidden Neurons - Airfoil}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_layers.png}
\caption{First Generation Median Training Error for variable number of
hidden Layers - Airfoil}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png}
\caption{First Generation final Training Error for variable number of
hidden Layers - Airfoil}
\end{figure}

@@

Regarding the hidden neurons, the dimension of each hidden layer, I used
two different strategies: Regular DAE-GP uses the same strategy as
described by earlier work (see {[}17{]} or {[}18{]}), the number of
hidden neurons is dynamically set at each generation to the maximum
length of all solutions inside the current training population. For
pre-trained DAE-GP another strategy had to be used, I used a fixed
number of hidden neurons for each run that is set equal to the maximum
length of all individuals that are present in the initial pre-training
population. I decided to take this aproach instead of using a fixed
number of hidden neurons for the fact that it simplifies the adjustment
of the parameter for later experiments on different problems. It should
nonetheless be mentioned that this strategy does give pre-training on
average more hidden neurons per layer resulting in a more complex
network (since the pre-training population is much larger than the
regular population, the chances for larger individuals inside it are
also higher).

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_median_fitness_byGens.png}
\caption{Best Fitness over 30 Generations - Airfoil}
\end{figure}

The result of the MWU test is 0

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/final_fit_boxplot.png}
\caption{Best Fitness Distribution after 30 Generations - Airfoil}
\end{figure}

\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_popDiversity_byGens.png}
Since {[}17{]} shows, that DAE-GP finds higher quality that are also
smaller in size than those found by average GP for a given number of
10.000 fitness evaluations, I also studied the differences in average
solution size between pre-trained and regular DAE-GP.

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=8.33333in]{./img/airfoil_FullRun_30gens/mean_Size_byGens.png}
\caption{Average Solution Size over 30 Generations - Airfoil}
\end{figure}
